|

### 2025-01-31 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
rerun models because of wrong priors, proofreading

---
title: >
  Preschoolers can form conventional pacts with each other 
  to communicate about novel referents
bibliography: preschool-tangrams.bib
csl: apa7.csl
document-params: "10pt, letterpaper"

@@

One influential early study suggested that 4-5-year-old preschoolers struggle with child-child referential communication [@glucksberg1966]. In their paradigm, one child was given a set of 6 blocks in a specific order. Their task was to describe the images on the block so their partner could pick out the corresponding block. As children described and selected blocks, they stacked them on pegs. While 4-5-year-old children succeeded on practice trials with familiar shapes and visual access to each other's blocks, children failed on critical trials where the blocks had abstract drawings and there was no visual access. Even after multiple rounds with the same images, children were not able to correctly order the blocks. @glucksberg1966 attributed children's communicative failures to their production of ego-centric descriptions that did not account for the other child's perspective.

Similar experiments with older children indicated a gradual improvement through adolescence both for initial accuracy and for the increase in accuracy across repetitions. Still, even the 9th grade sample was noticeably worse than the adult college student sample [@glucksberg1967; @krauss1969]. Given that even teenagers had difficulties with the task, the complex stacking paradigm and the large number of potential targets may have posed task demands preventing accurate measurement of children's abilities.

@@

More recently, though no evidence has directly contradicted this early study, a number of other studies have revealed early emerging skills in the preschool years that support the use of referential communication. Preschool-aged children are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010] and 6-year-olds are more likely to use consistent referential expressions when their partner is consistent [@koymen2014], suggesting that young children are sensitive to the norm of consistent descriptions in cooperative communication. Further, preschool-aged children adapt the informativeness of their referential expressions based on the visual content available to their interlocutor [@matthews2006;@nadig2002;@nilsen2009]. By age 5, children integrate information about other's perspectives into their comprehension of utterances [@sanjuan2015]. Overall, by late preschool, children can can reason about others’ perspectives to communicate more effectively.

@@

Our goal for Experiment 1 was to test young children's ability to coordinate on descriptions of abstract shapes that their partner could understand. Young children can be very sensitive to task demands and cognitive load [@turan-kucuk2024; @keen2003; @carruthers2013], so we adapted the experimental framework from @leung2024, and further simplified it by reducing the total pool of targets and the number of trials. This experiment was pre-registered at https://osf.io/kcv8j?view_only=fd55733671194962b868a4ba19377631.

@@

Children completed two warm-up trials with black and white images of familiar shapes, followed by 3 blocks of the 4 target images (Figure \ref{game}C). Targets were randomly paired with another of the critical images as the foil. 
The experimenters running the game did not volunteer descriptions, but they did scaffold the interaction, prompting children to describe the images, and sometimes repeating children's statements (especially when utterances were inaudible or the child did not respond immediately; this aspect of the procedure was modified in Experiment 2). The entire interaction was video-recorded.

@@

Statistical analyses were run in brms [@burkner2018] with weakly informative priors. We present the estimate and 95% credible intervals. The experimental set-up, analysis code, de-identified transcripts, and performance data for both experiments is available at \url{https://osf.io/7d3ne/?view_only=d094ccd6438741bfbcaa7fde9fbb56b2}.

@@

```{=latex}
\begin{table}
	\caption{Example descriptions children successfully used used to identify different target images in Experiment 2.}
	\label{example}
	\begin{tabular}[t]{p{16em}p{4em}}
		\hline
		%hold
		$\bullet$ person & 		\multirow{6}{*}{\includegraphics[width=.5in]{I1.jpg}}\\

@@

As Experiment 2 was very similar to Experiment 1, we focus on the changes made compared to Experiment 1. Experiment 2 was pre-registered at https://osf.io/y2dax/?view_only=fd55733671194962b868a4ba19377631.

@@

As Experiment 2 had 4 blocks, we examined whether descriptions were converging semantically toward the final description. We compared the utterances from the first three blocks to the descriptions in the last block using a Bayesian mixed effects linear regression predicting similarity.^[`r sim_to_last_form |> form()`] Over the first three blocks, descriptions became increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`). Descriptions were more similar to each other if they came from the same child, which is expected as a sign of internal consistency (`r sim_to_last_2_results |> stats_text(3,3)`). Although over time descriptions did get more similar to the last block utterance, the semantic distance between adjacent block utterances was relatively constant (`r sim_to_next_2_results |> stats_text(2,3)`).

Partnerships often diverged from one another as groups focused on distinct aspects of the image. We tested whether descriptions in different games to the same target diverged over time using a Bayesian mixed effects linear regression.^[`r sim_across_form |> form()`] As the games progressed, descriptions from different games became slightly further apart in semantic space (`r sim_across_2_results |> stats_text(2,3)`), which indicates that games are converging to different conventions. These patterns of increasing similarity within games and increasing divergence between games qualitatively match the patterns found for adults [@boyce2024].

```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between descriptions from earlier blocks (1-3) and the last block.  Heavy dots are means with bootstrapped 95% CIs; light dots are individual values. \\label{tolast}" }
sims_2 |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == 4) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "Same teller", "Different teller")) |>
  ggplot(aes(earlier, sim, color = target1)) +
    geom_smooth(color="black", method="lm")+
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
    #stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2), geom="line", lty="dashed") +
  #  geom_smooth( method="lm")+
  facet_wrap(~same_speaker) +
  scale_color_solarized() +
  scale_x_continuous(breaks=c(1,2,3), labels=c("1", "2", "3"))+
  labs(x = "Comparing an earlier block (1-3) to last block (4)", y = "Cosine Similarity", color = "Target") +
  theme(legend.position = "bottom",
            strip.background = element_blank(),
        #panel.grid=element_blank()
)

@@

We might expect that whether a description was successful influences whether the same description, or a variant of it, is employed in future rounds. Intuitively, successful descriptions can be copied and built upon, while unsuccessful descriptions should be replaced by a fresh attempt. To test this idea that accuracy might be predictive of similarity to future descriptions, we ran a post-hoc Bayesian linear model predicting similarity to the next block description in terms of accuracy.^[`r form(sim_acc_to_next_form)`] Descriptions that elicited a correct response were more similar to the next block description (`r stats_text(sim_acc_to_next_mod_meta, 2, 3)`) with no substantial interaction with block number or whether both descriptions came from the same teller. This pattern of results is consistent with the expectation that children are more likely to stick to their own description or repeat the other child's description if it was previously successful. <!--Numerically, the size of the boost for being correct was similar to the size of the boost from coming from the same person `r stats_text(sim_acc_to_next_mod_meta, 6,3)`. -->

@@

Prominent early findings have been interpreted as evidence that young children cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966]. However, more recent developmental work has found that young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low. <!--Recent work has established that children are successful at producing and understanding referential expressions to novel shapes when playing reference games with their parents. -->We revisited the question of preschoolers' performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.

@@

It is unclear to what extent the uniformly short descriptions we observed are a product of the simplified task or children's behavioral differences from adults. In this case, the low number of options and relatively easy to describe shapes may have obviated the need for long initial descriptions. Indeed, adult controls in @leung2024 used shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images [@hawkins2020b; @boyce2024]. However, young children may also struggle to produce longer descriptions, and young children may be more willing to take guesses when adults would seek additional clarification. Especially in light of other work suggesting that conceptual pact formation and reduction in utterance length sometimes decouple in adults [@boyce2024], further empirical work on the factors driving verbosity in reference games is warranted.

>>>>

---
title: |
  | Preschoolers can form conventional pacts with each other 
  | to communicate about novel referents
bibliography: preschool-tangrams.bib
csl: apa7.csl
document-params: "10pt, letterpaper"

@@

One influential early study suggested that 4-5-year-old preschoolers struggle with child-child referential communication [@glucksberg1966]. In their paradigm, one child was given a set of 6 blocks in a specific order. Their task was to describe the image on each block so their partner could pick out their corresponding block. As children described and selected blocks, they stacked them on pegs. While 4-5-year-old children succeeded on practice trials with familiar shapes and visual access to each other's blocks, children failed on critical trials where the blocks had abstract drawings and there was no visual access. Even after multiple rounds with the same images, children were not able to correctly order the blocks. @glucksberg1966 attributed children's communicative failures to their production of ego-centric descriptions that did not account for the other child's perspective.

Similar experiments with older children indicated a gradual improvement through adolescence both for initial accuracy and for the increase in accuracy across repetitions. Still, even the 9th grade sample was noticeably worse than the adult college student sample [@glucksberg1967; @krauss1969]. Given that even teenagers had difficulties with the task, the complex stacking paradigm and the large number of potential targets may have posed task demands that prevented accurate measurement of children's abilities.

@@

More recently, though no evidence has directly contradicted @glucksberg1966, a number of other studies have revealed early emerging skills in the preschool years that support the use of referential communication. Preschool-aged children are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010] and 6-year-olds are more likely to use consistent referential expressions when their partner is consistent [@koymen2014], suggesting that young children are sensitive to the norm of consistent descriptions in cooperative communication. Further, preschool-aged children adapt the informativeness of their referential expressions based on the visual content available to their interlocutor [@matthews2006;@nadig2002;@nilsen2009]. By age 5, children integrate information about other's perspectives into their comprehension of utterances [@sanjuan2015]. Overall, by late preschool, children can can reason about others’ perspectives to communicate more effectively.

@@

Our goal for Experiment 1 was to test young children's ability to coordinate to produce descriptions of abstract shapes that their partner could understand. Young children can be very sensitive to task demands and cognitive load [@turan-kucuk2024; @keen2003; @carruthers2013], so we adapted the experimental framework from @leung2024, and further simplified it by reducing the total pool of targets and the number of trials. This experiment was pre-registered at [anonymized link](https://osf.io/kcv8j?view_only=fd55733671194962b868a4ba19377631).

@@

Children completed two warm-up trials with black and white images of familiar shapes, followed by 3 blocks of the 4 target images (Figure \ref{game}C). Targets were randomly paired with another of the critical images as the foil.

The experimenters running the game did not volunteer descriptions, but they did scaffold the interaction, prompting children to describe the images, and sometimes repeating children's statements (especially when utterances were inaudible or the child did not respond immediately; this aspect of the procedure was modified in Experiment 2). The entire interaction was video-recorded.

@@

Statistical analyses were run in brms [@burkner2018] with weakly informative priors. We report estimates and 95% credible intervals. The experimental set-up, analysis code, de-identified transcripts, and performance data for both experiments is available at [anonymized repo](https://osf.io/7d3ne/?view_only=d094ccd6438741bfbcaa7fde9fbb56b2).

@@

```{=latex}
\begin{table}
	\caption{Example descriptions children successfully used to identify different target images in Experiment 2.}
	\label{example}
	\begin{tabular}[t]{p{16em}p{4em}}
		\hline
		%hold
		$\bullet$ person & 		\multirow{6}{*}{\includegraphics[width=.5in]{I1.jpg}}\\

@@

As Experiment 2 was very similar to Experiment 1, we focus on the changes made compared to Experiment 1. Experiment 2 was pre-registered at [anonymized link](https://osf.io/y2dax/?view_only=fd55733671194962b868a4ba19377631).

@@

As Experiment 2 had 4 blocks, we examined whether descriptions were converging semantically toward the final description. We compared the utterances from the first three blocks to the descriptions in the last block using a Bayesian mixed effects linear regression predicting similarity.^[`r sim_to_last_form |> form()`] Over the first three blocks, descriptions became increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`). Descriptions were more similar if they came from the same child, which is expected as a sign of internal consistency (`r sim_to_last_2_results |> stats_text(3,3)`). Although over time descriptions did get more similar to the last block utterance, the semantic distance between adjacent block utterances was relatively constant (`r sim_to_next_2_results |> stats_text(2,3)`).

Partnerships often diverged from one another as groups focused on distinct aspects of the image. We tested whether descriptions in different games diverged over time using a Bayesian mixed effects linear regression.^[`r sim_across_form |> form()`] As the games progressed, descriptions to the same target from different games became slightly less similar (`r sim_across_2_results |> stats_text(2,3)`), which indicates that games are converging to different conventions. These patterns of increasing similarity within games and increasing divergence between games qualitatively match the patterns found for adults [@boyce2024].

```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between descriptions from earlier blocks (1-3) and the last block in Experiment 2.  Heavy dots are means with bootstrapped 95% CIs; light dots are individual values. \\label{tolast}" }
sims_2 |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == 4) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "Same teller", "Different teller")) |>
  ggplot(aes(earlier, sim, color = target1)) +
    geom_smooth(color="black", method="lm")+
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
    #stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2), geom="line", lty="dashed") +
  #  geom_smooth( method="lm")+
  facet_wrap(~same_speaker) +
  scale_color_solarized() +
  scale_x_continuous(breaks=c(1,2,3), labels=c("1", "2", "3"))+
  labs(x = "Comparing an earlier block (1-3) to last block (4)", y = "Cosine Similarity", color = "Target") +
  theme(legend.position = "bottom",
            strip.background = element_blank(),
        #panel.grid=element_blank()
)

@@

We might expect that whether a description was successful influences whether the same description, or a variant of it, is employed in future rounds. Intuitively, successful descriptions can be copied and built upon, while unsuccessful descriptions should be replaced by a fresh attempt. To test whether accuracy is predictive of similarity to future descriptions, we ran a post-hoc Bayesian linear model predicting similarity to the next block description in terms of accuracy.^[`r form(sim_acc_to_next_form)`] Descriptions that elicited a correct response were more similar to the next block description (`r stats_text(sim_acc_to_next_mod_meta, 2, 3)`) with no substantial interaction with block number or whether both descriptions came from the same teller. This pattern of results is consistent with the expectation that children are more likely to stick to their own description or repeat the other child's description if it was previously successful. <!--Numerically, the size of the boost for being correct was similar to the size of the boost from coming from the same person `r stats_text(sim_acc_to_next_mod_meta, 6,3)`. -->

@@

Prominent early studies claimed that young children cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966]. However, more recent developmental work has found that young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low. <!--Recent work has established that children are successful at producing and understanding referential expressions to novel shapes when playing reference games with their parents. -->Here, we revisited the question of preschoolers' performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.

@@

It is unclear to what extent the uniformly short descriptions we observed are a product of the simplified task or children's behavioral differences from adults. In this case, the low number of options and relatively easy-to-describe shapes may have obviated the need for long initial descriptions. Indeed, adult controls in @leung2024 used shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images [@hawkins2020b; @boyce2024]. However, young children may also struggle to produce longer descriptions, and young children may be more willing to take guesses when adults would seek additional clarification. Especially in light of other work suggesting that conceptual pact formation and reduction in utterance length sometimes decouple in adults [@boyce2024], further empirical work on the factors driving verbosity in reference games is warranted.
